---
title: "Regression Models"
author: "Patricio Del Boca"
date: "24/10/2015"
output: pdf_document
---
# Executive Summary

You work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome).

They are particularly interested in the following two questions:

 - Is an automatic or manual transmission better for MPG
 - Quantify the MPG difference between automatic and manual transmissions

## Loading Data
```{r, results='hold'}
data(mtcars)
mtcars$cyl <- factor(mtcars$cyl)
mtcars$vs <- factor(mtcars$vs)
mtcars$am <- factor(mtcars$am)
levels(mtcars$am) <- c("Automatic", "Manual")
mtcars$gear <- factor(mtcars$gear)
mtcars$carb <- factor(mtcars$carb)
head(mtcars)
```

#### Summary exploration of MPG and Transmition
```{r}
summary(mtcars$mpg)
table(mtcars$am)
```

## Data Analysis
For code reference got to the **Appendix Section**, Plot 1.

```{r, echo=FALSE}
library(ggplot2)
g <- ggplot(mtcars, aes(x=am,y=mpg, fill=am)) + geom_boxplot()
g <- g + xlab("Transmission") + ylab("Miles per Galon")
g <- g + ggtitle("Miles Per Galon by Transmission \n (Plot 1)")
g
```

At first glance, it seems that Manual Transmission cars have higher MPG than 
Automatic cars.

Let's try to quantify it.

## Linear Model
```{r}
fit <- lm(mpg ~ am, data = mtcars)
summary(fit)$coefficients
```
#### Conclussions:
 - The **average of Automatinc Transmissions MPG** is 17.14.
 - The **difference in the empirical mean between Manual and Automatic MPG** is 7.24.
 - Is Automatic different to Manual? The 0.0002 p-value (less than 0.05) is actually saying that **there is a statistical difference in MPG between cars with Automatic and Manual Transmission**.

Lets see how well the Transmission explains the variability in MPG:

```{r}
summary(fit)$r.squared
```

Only 36% of the variability in MPG values is explained by the Transmission. Lets
fit some models with more variables to try to explain more

## Multivariate Linear Model

```{r}
fullModel <- lm(mpg ~ ., data=mtcars)
summary(fullModel)$coefficients
```

##### How do we interpret the p-values?

- A predictor that has a **low p-value** is likely to be a meaningful addition to our model because changes in the predictor's value are related to changes in
the response variable.
 - A **larger p-value** suggests that changes in the predictor are
not associated with changes in the response.

There aren't good predictors since none of the p-values are lower than 0.05.

Let's simplify the model by reducing it's variables:

```{r, results='hide'}
stepModel <- step(fullModel, k=log(nrow(mtcars)))
```
```{r, results='hold'}
summary(stepModel)$coefficients
summary(stepModel)$r.squared
```


Much better! This variables (wt + qsec + am) can explain the 85% of the 
variability in MPG values. Also, all the coefficients are statistically 
signifficant and also suggest that **wt** is the most explanatory variable.

## Apendix

#### Data Analysis Plots

##### Plot 1
```{r, results='hide'}
library(ggplot2)
g <- ggplot(mtcars, aes(x=am,y=mpg, fill=am)) + geom_boxplot()
g <- g + xlab("Transmission") + ylab("Miles per Galon")
g <- g + ggtitle("Miles Per Galon by Transmission \n (Plot 1)")
g
```

```{r}
plot(stepModel)
```
 - Residuals vs. Fitted: Shows no consistent pattern, it supports the accuracy of the independence assumption. Also, no other patterns are recognized (exponential, sinusoidal) son no further exploration is suggested
 - Normal Q-Q: The residuals are normally distributed because the points are closely to the line. the further from the mean, the more distributed the points are.
 - Scale-Location: confirms the constant variance assumption, as the points are randomly distributed.
 - Residuals vs. Leverage: that no outliers are present, as all values fall well within the 0.5 bands. 

