---
title: "Regression Models"
author: "Patricio Del Boca"
date: "24/10/2015"
output: pdf_document
---
# Executive Summary

The objective of this document is looking at a data set of a collection of cars, and explore the relationship between a set of variables and miles per gallon (MPG) (outcome).

There is a particular interest in the following two questions:

 - Is an automatic or manual transmission better for MPG
 - Quantify the MPG difference between automatic and manual transmissions

Key takeaways:

 - Regarding the first question: Manual Transmissions are better for MPG.
 - Regarding the second question: Manual transmission are better by a factor of 1.8. Also, the means of MPG are better by 7.25.

## Data Analysis

At first glance, it seems that Manual Transmission cars have higher MPG than 
Automatic cars. Let's try to quantify it. For code reference and graphics got to the **Appendix Section**, Plot 1.

## Linear Model
```{r, results='hold'}
fit <- lm(mpg ~ am, data = mtcars)
summary(fit)$coefficients
summary(fit)$r.squared
```

#### Conclussions:
 - The **average of Automatinc Transmissions MPG** is 17.14.
 - The **difference in the empirical mean between Manual and Automatic MPG** is 7.24.
 - Is Automatic different from Manual? The 0.0002 p-value (less than 0.05) is actually saying that **there is a statistical difference in MPG between cars with Automatic and Manual Transmission**.
 -  The R Squared value tells us that **only 36% of the variability in MPG values is explained by the Transmission**.

\newpage

## Multivariate Linear Model

```{r, results='hold'}
fullModel <- lm(mpg ~ ., data=mtcars)
summary(fullModel)$adj.r.squared
sum(summary(fullModel)$coef[,4] < 0.05)
```

Although the Adjusted R-squared is 0.77, there aren't good predictors since none of the p-values are lower than 0.05. For more details on coefficients check **Appendix: Full Model Coefficients**.

```{r, results='hide'}
stepModel <- step(fullModel, k=log(nrow(mtcars)))
```
```{r, results='hold'}
summary(stepModel)$coefficients
summary(stepModel)$r.squared
```

Much better! The variables **(wt + qsec + am)** can explain the 85% of the 
variability in MPG values. Also, all the coefficients are statistically 
signifficant and also suggest that **wt** is the most explanatory variable.

## Residuals Analysis
For details on the plot check **Apendix: Residuals Plot**.

- **Residuals vs. Fitted:** Shows no consistent pattern, it supports the accuracy of the independence assumption. Also, no other patterns are recognized (exponential, sinusoidal) son no further exploration is suggested
- **Normal Q-Q:** The residuals are normally distributed because the points are closely to the line. the further from the mean, the more distributed the points are.
- **Scale-Location:** confirms the constant variance assumption, as the points are randomly distributed.
- **Residuals vs. Leverage:** that no outliers are present, as all values fall well within the 0.5 bands. 

## Inference

Assuming that cars with Manual and Automatic Transmission are from the same population as the Null Hypothesis, the p-value of 0.00137 reject it. So, the automatic and manual transmissions are from different populations.

\newpage

## Apendix 

Data Loading and Summarization:

```{r, results='hold'}
data(mtcars)
mtcars$cyl <- factor(mtcars$cyl)
mtcars$vs <- factor(mtcars$vs)
mtcars$am <- factor(mtcars$am)
levels(mtcars$am) <- c("Automatic", "Manual")
mtcars$gear <- factor(mtcars$gear)
mtcars$carb <- factor(mtcars$carb)
summary(mtcars$mpg)
table(mtcars$am)
```

Plot 1: MPG by Transmission Boxplot

```{r}
library(ggplot2)
g <- ggplot(mtcars, aes(x=am,y=mpg, fill=am)) + geom_boxplot()
g <- g + xlab("Transmission") + ylab("Miles per Galon")
g <- g + ggtitle("Miles Per Galon by Transmission \n (Plot 1)")
g
```

Full Model Coeficients

```{r}
summary(fullModel)$coef
```

How do we interpret the p-values?

- A predictor that has a **low p-value** is likely to be a meaningful addition to our model because changes in the predictor's value are related to changes in
the response variable.
 - A **larger p-value** suggests that changes in the predictor are
not associated with changes in the response.

\newpage
Residuals Plot:
```{r}
plot(stepModel)
```

Inference
```{r}
result <- t.test(mtcars$mpg ~ mtcars$am)
result$p.value
result$estimate
```